Lesson Name:  Basic Operations
Course Name:  Introduction To R
Type:         Standard
Author:       VP Nagraj & Stephen Turner
Organization: University of Virginia
Version:      2.3.1
================================================================

--- &text

Welcome to the 'Introdcution To R' swirl course. The first lesson ('Basic Operations') will cover how to use some basic math functions and assignment of values to objects.  Let's get going.

--- &text

We'll start by learning about RStudio. R is the underlying statistical computing environment, but using R alone is no fun. RStudio is a graphical integrated development environment that makes using R much easier.

--- &text

The RStudio interface is a window divided into panes - these can include an editor, console, environment/history and plots/help. If you're reading this then you're looking at the console pane. You can edit this layout by going to 'Tools' > 'Global Options' > 'Pane Layout'. Now that we know a little about RStudio we can dive into R. 

--- &cmd_question

R can be used as a glorified calculator. Basic math operations like addition, subtraction, multiplication, etc. all have their own operators. For example typing 2 * 7 in the console will give you 14 ... 2 + 7 will give you 9 ... and so on. Let's dividing 2 by 7 and see what we get.

```{r}
2 / 7 
```

*** .hint
The syntax for a division operation is simply x / y where x is the numerator and y is the denominator.

--- &text
R is also smart enough to understand order of operations. For example 2 + 3 * 4/(5 + 3) will be 3.5

--- &mult_question

What is 5 ^ 2 + 1 according to R?

_1. 26_
2. 125

*** .hint
Without parentheses to slow it down R will walk through the operations one at a time.

--- &text

R works through the order of operations ... so without parentheses it will evaluate each operator in sequence. That's great. But so what?  

--- &text

To be able to do truly interesting things in R we need to assign values to objects. To create an object, we give it a name followed by the assignment operator <- and its value. For example if we were recording Michael Jordan's height in centimeters we could store it as something like: height_cm <- 200

--- &text
It's worth emphasizing that <- is the operator that assigns values on the right to objects on the left. You can think of it like an arrow that points from the value to the object. This operator is mostly similar to = but not always. Learn to use <- as it is good programming practice. Using = in place of <- can lead to issues down the line.

--- &text

Objects can be given any name such as x, current_temperature or subject_id. You want your object names to be explicit and not too long. They cannot start with a number (2x is not valid but x2 is). R is case sensitive and not going to forgive your spelling errors - weight_kg is different from Weight_kg is different from wieght_kg, etc.

--- &text
Note that there are some names that cannot be used because they represent the names of fundamental functions in R (e.g., if, else, for). In general, even if it's allowed, it's best to not use other function names, which we'll get into shortly (e.g., c, T, mean, data, df, weights). If in doubt check the help to see if the name is already in use. RStudio will also autocomplete to show you if an object (or function) with that name already exits. It's also best to avoid dots (.) within a variable name as in my.dataset. It is also recommended to use nouns for variable names, and verbs for function names.

--- &cmd_question

Okay ... let's try it out. Create an object that's called weight_kg and has a value of 90

```{r}
weight_kg <- 90 
```

*** .hint
Make sure you are using the <- assignment operator and spelling the weight_kg variable name correctly, and with the underscore and no spaces.

--- &cmd_question

When assigning a value to an object, R does not print anything. You can force to print the value by typing the name. Type weight_kg and see how it works.

```{r}
weight_kg
```

*** .hint
To print the value of an object just type its name ... in this case weight_kg

--- &cmd_question
Now that R has weight_kg in memory, we can do arithmetic with it. For instance, we may want to convert this weight in pounds. We don't want to lose that conversion so let's store it in a new variable called weight_lb. Give it a shot ... and in case you're rusty on metric conversion, weight in pounds is 2.2 times the weight in kg.

```{r}
weight_lb <- 2.2 * weight_kg
```

*** .hint
To create a new object, type its name (weight_lb) followed by the assignment operator and then the function that yields its value. In this example use 2.2 then the multiplication symbol (*) and the first object's name (weight_kg)

*** .ans_tests
any_of_exprs('weight_lb <- 2.2 * weight_kg', 'weight_lb <- weight_kg * 2.2')

--- &cmd_question
We can also change a variable's value by assigning it a new one. Let's say for instance the weight_kg is now 100. All we need to do is use the same assignment syntax with the new value. Go for it.

```{r}
weight_kg <- 100
```

*** .hint
Use the <- operator with the same object name as before (weight_kg) but a new value (100).

--- &mult_question

Now that you've changed weight_kg to 100, if you were to print the weight_lb variable to the console what would its value be?

_1. 198_
2. 220

*** .hint
Even though you've updated weight_kg you haven't updated weight_lb from its original assignment.

--- &text

It's important to understand that assigning a value to one variable does not change the values of other variables. So in this case if you wanted to update the weight_lb variable so that it's value was 220 then you would have to reassign it again after setting weight_kg to 100.

--- &text
You can see what objects (variables) are stored by viewing the Environment tab in Rstudio. You can also use the ls() function. You can remove objects (variables) with the rm() function. You can do this one at a time or remove several objects at once.

--- &cmd_question
Things were getting a little confusing with the converted weight_lb variable. Go ahead and remove it.

```{r}
rm(weight_lb)
```

*** .hint
Use the rm([INSERT OBJECT NAME HERE]) to remove an object.

--- &text
Good job. You've made it through the first lesson. Next up is 'Functions'.

